Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-02-18 Nicolas.tex
Finished YC changes (at pdf/hist and BBC hist près)

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see there that the redshift limits we have selected roughly
correspond to the peak of these histograms. \nn{As the peaks indicate that we
are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
cuts fall near said peaks.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the Pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
    \label{fig:sample}
\end{figure}

@@

\nn{In~\cite{rigault2018}, we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. Given the structure of the
stretch-LsSFR scatter plot, our model is defined as follows: 
\begin{itemize}
    \item for the young population the underlying stretch distribution
    is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
    \item the old population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
    \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
    same as the young population one. \yc{explain meaning of $a$}
\end{itemize}
\nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1(z)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1(z) = {}}
    \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
    (1 - \delta(z)) &\times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
\end{align}

The \yc{maximum-likelihood estimate} of the 5 free parameters
$\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
minimizing a pseudo-$\chi^2$ $\mathcal{L}$ defined as:
\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}},
\end{equation}
with
\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
    y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
        (1-y^i) &\times \nn{\bigg[} 
        a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
     &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \nn{\bigg]}
\end{align}
where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error 
and the probability that the SN is young, respectively. 
In practice, to ensure fraction $a$ is constrained between 0 and 1,
we fit for $\alpha$ such
that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.

@@

\cite{howell2007} used a simpler model with a normal distribution for each young 
and old populations. \nn{We thus consider a ``Howell+drift'' model, with one 
single Gaussian per age group}. \yc{one has to explain the ``drift'' part!!!}

\yc{Alternatively, since we aim at probing the existence of a redshift drift, 
we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{fix labels!!!}

@@

We applied each model on both the reference and conservative samples (cf.
section~\ref{sec:sample}). Because the models presented
in section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This estimator
penalizes extra degrees of freedom to avoid over-fitting the data. It is defined
as follow:
\begin{equation}
    \mathrm{AIC} = \yc{\mathcal{L}_{min}} + 2k
\end{equation}
where \yc{$\mathcal{L}_{min}$} is
the minimum value of the pseudo-$\chi^2$ as defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The \nn{reference} model
is the one with the smallest AIC, and the probability for another model to be at
least as representative of the data as this one is given by:
\begin{equation}
    p(\mathrm{other} \geq \mathrm{ref}) =
    \exp\left(\Delta\mathrm{AIC}/2\right)
\end{equation}
\yc{with $\Delta\mathrm{AIC} < 0$}. Results are presented in Table~\ref{tab:comp} and are illustrated
Fig.~\ref{fig:mod_comp}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
    \caption{$\Delta$AIC between the reference ``Base'' model and the other
    models (see also Table~\ref{tab:comp}). The full and open blue markers
correspond to model with and without redshift drift, respectively. Light markers
show the results when this analysis is performed on the conservative sample 
rather than the fiducial one. Color-bands illustrate the validity of the models, from 
Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded ($\Delta\mathrm{AIC} < -20$),
see text \yc{This is not explained in the text!}. \yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
    \label{fig:mod_comp}
\end{figure}

@@

\mr{\yc{Furthermore,} we find that underlying stretch distributions not including 
    a redshift drift are all
    excluded to describe the data as well as the Base model. In fact,
    the best non drifting model (Asymmetric) has \yc{a very marginal chance 
    ($p = 2\times10^{-6}$)} \yc{not same number as in table!!!} to
    describe the data as well as the Base model.}
    \yc{This result is just a quantitative assessment of qualitative facts clearly visible in 
    Fig.~\ref{fig:modelall}: the mean SN stretch per bin of redshift strongly 
    suggests a redshift evolution rather than having a constant value, and this evolution is compatible with the redshift drift expected from Eq.~\ref{eq:delta}.}

\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
fig. Mickael, please rephrase and remove what is not needed anymore.}

@@

\yc{This paragraph is not clear: what is the impact, and its importance, on which case? Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}
\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
    been explicitly accounted for in cosmological analyses. Not doing so is a
    second order issue for SN cosmology as it only affects the way one does
    Malmquist bias correction. Indeed, as long as the Phillip's relation
    standardization parameter $\alpha$ is not redshift dependent (a study behind
    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
    for cosmology are blind to the underlying stretch distribution.  However,
    since the first missed SNe~Ia are the faintest, which are also those with
    the lowest stretch, once we include SNe~Ia affected by selection effects we
    need to account for the fact that only the brightest SNe~Ia exist in the
    low-stretch end of the sample after standardization. The modeling of the SN
    stretch then enters the analysis to understand which SNe~Ia are missing
given the selection functions to mimic a selection-bias free sample.}

@@

\mr{The state-of-the-art tool for doing so \yc{what?} is the BBC described in
    \cite{scolnic2016} and \cite{kessler2017}, which is used in recent SN
    cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
    descosmopaper2019} including the direct measurement of H$_0$
\citep{riess2016,riess2019}.}

\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation \yc{in our analysis}. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the smallest of all
    non-drift models (but still $\Delta \mathcal{L}=-11.5$ worse than the
    reference Base model), it is penalized by requiring 15 free parameters (3 for each of the 5 samples of the analysis).
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because \yc{traditional} surveys span limited redshift ranges, the
    per sample approach can somewhat account for \yc{implicit} redshift drifts. We
    stress here that, as measurements of modern surveys cover \yc{increasingly} larger redshift
    ranges to reduce calibration systematic  uncertainties, this approach is becoming
    less \yc{valid}, notably for PS1, DES and, soon, LSST \yc{redshift ranges? Why not SNLS too?}. Already with the current
    surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
    representation of the data as our best model ($p=4\times 10^{-8}$). A more
    detailed discussion of the consequence of this results for cosmology follows
in section~\ref{sec:bbc}.}

\mr{We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and
    $\mu_0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
    of \nn{a} selection-bias free sample, see
    section~\ref{sec:sample}. \yc{WHY??? The developed BBC for \textit{incomplete} samples!} If we were to use \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go even deeper from $-32$ to $-47$. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. \yc{???} \yc{Last},
    fixing the Base model parameters to those derived using SNf data only (top
    row Table~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$, but removing the SNf sample reduces this difference to
    $\Delta\mathrm{AIC}=-5$ \yc{not clear: the SNf sample is removed or not???}. Such a \yc{dramatic apparent improvement} is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift \nn{lever} arm to probe the impact of redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability $y^i$ for a SN to be young is
    known ($y^i = p_y{}^i$) while other samples rely on a probabilistic proxy for
    $y^i$, namely the fraction $\delta(z^i)$ 
    of young SNe~Ia at SN redshift. While in any case the Base model is favored, the relative weight
    of the SNf sample highlights the importance of adding non-targeted nearby
surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
graham2019} to unambiguously confirm our \nn{model}.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        distributions for the PS1 sample. The Green line (band) shows our best
        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
        approach. The dashed line shows the noe \yc{???} estimated by
    \cite{scolnic2018a}. The blue to red curves display our Base drifting
stretch model estimated at each PS SN redshift; bluer for lower redshifts.}
    \label{fig:bbc_pdf_ps1}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
    \caption{Difference of mean stretch standardization (in mag) between the BBC
        algorithm used in \cite{scolnic2018a} and our Base drifting model.
        \nn{The small colored points show the SDSS, PS1 and SNLS evolution,
            while the orange markers represent their mean evolution in 30
            regular redshift bins.} For comparison, we show in gray bands the
            expected magnitude difference between $\Lambda$CDM and $w$CDM
            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in \yc{blue},
            more exotic $w_0,w_a$CDM examples. This figure illustrates that the
            amplitude of the potential magnitude bias caused by inaccurate
        modeling of the underlying stretch distribution is comparable to the
    observational effect of exotic form of dark energy. \yc{Is this fig. still relevant?}}
    \label{fig:magdrift}
\end{figure}

@@

We have presented a study of the drift of the underlying \yc{I still prefer ``intrinsic''} SNe~Ia stretch
distribution as a function of redshift. We used SNe~Ia from magnitude-limited
surveys from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
well as HST, to which we added SNfactory data from \cite{rigault2018} for the
low-redshift bin. We only considered the supernovae that have been discovered
in redshift ranges of each surveys where selection effect are negligible. This
way the observed SNe~Ia stretch are random sampling of the true underlying
distribution. This corresponds to 569 SNe~Ia (422 when considering more
conservative cuts).

>>>>

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see there that the redshift limits we have selected roughly \nn{but logically}
correspond to the peak of these histograms. \nn{as the peaks indicate that we
are starting to miss the faintest SNe~Ia}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-inferno.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the Pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-inferno.pdf}
    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
    \label{fig:sample}
\end{figure}

@@

\nn{In~\cite{rigault2018}, we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. Given the structure of the
stretch-LsSFR scatter plot, our model is defined as follows: 
\begin{itemize}
    \item for the young \nn{(i.e. $\log \mathrm{LsSFR}>-10.82$)} population, the underlying stretch distribution
    is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
    \item the old \nn{(i.e. $\log \mathrm{LsSFR}<-10.82$)} population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
    \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
    same as the young population one\nn{, $a$ representing the relative importance
    of the two Gaussians}.
\end{itemize}
\nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1(z)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1(z) = {}}
    \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
    (1 - \delta(z)) &\times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
\end{align}

The \yc{maximum-likelihood estimate} of the 5 free parameters
$\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
minimizing a pseudo-$\chi^2$ $\mathcal{L}$ defined as:
\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}},
\end{equation}
with
\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
    y^i &\times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} + {} \nonumber\\
        (1-y^i) &\times \nn{\bigg[} 
        a \times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_1,
        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} + {} \nonumber\\
     &(1-a) \times \yc{\mathcal{N}\left(x^{i}_{1} \mid \mu_2,
     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right)} \nn{\bigg]}
\end{align}
where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error 
and the probability that the SN is young, respectively. 
In practice, to ensure fraction $a$ is constrained between 0 and 1,
we fit for $\alpha$ such
that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.

@@

\cite{howell2007} used a simpler model with a normal distribution for each young 
and old populations. \nn{We thus consider a ``Howell+drift'' model, with one 
single Gaussian per age group and the $\delta(z)$ drift from \cite{rigault2018}}. \uline{\yc{one has to explain the ``drift'' part!!!} \nn{done}}

\yc{Alternatively, since we aim at probing the existence of a redshift drift, 
we are also testing non-drifting models by restricting the ``Base'' and ``Howell'' models to use a supposedly redshift-independent fraction $\delta(z) = f$ of young SNe; these models are labeled ``Base constant'' and ``Howell constant''.} \yc{\uline{fix labels!!! \nn{done}}}

@@

We applied each model on both the reference and conservative samples (cf.
section~\ref{sec:sample}). Because the models presented
in section~\ref{sec:modeling} have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare them. This estimator
penalizes extra degrees of freedom to avoid over-fitting the data. It is defined
as follow:
\begin{equation}
    \mathrm{AIC} = \yc{\mathcal{L}_{min}} + 2k
\end{equation}
where \yc{$\mathcal{L}_{min}$} is
the minimum value of the pseudo-$\chi^2$ as defined Eq.~\eqref{eq:likelihood} and $k$ is the number of free parameters. The \nn{reference} model
is the one with the smallest AIC, and the probability for another model to be at
least as representative of the data as this one is given by:
\begin{equation}
    p(\mathrm{other} \geq \mathrm{ref}) =
    \exp\left(\Delta\mathrm{AIC}/2\right)
\end{equation}
\yc{with $\Delta\mathrm{AIC} < 0$}. Results are presented in Table~\ref{tab:comp} and are illustrated Fig.~\ref{fig:mod_comp}\nn{, where a $\Delta$AIC of -5 corresponds to a $z$-score of 1.4, -10 to a $z$-score of 2.5, and -20 to a $z$-score of 3.9}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/mod_comp.pdf}
    \caption{$\Delta$AIC between the reference ``Base'' model and the other
    models (see also Table~\ref{tab:comp}). The full and open blue markers
correspond to model with and without redshift drift, respectively. Light markers
show the results when this analysis is performed on the conservative sample 
rather than the fiducial one. Color-bands illustrate the validity of the models, from 
Acceptable ($\Delta\mathrm{AIC} > -5$) to Excluded ($\Delta\mathrm{AIC} < -20$),
see text \yc{\uline{This is not explained in the text! \nn{Now done (conversion to z-score)}}}. \yc{According to the AIC}, all non-drifting models (open symbols) are excluded for the reference sample.}
    \label{fig:mod_comp}
\end{figure}

@@

\mr{\yc{Furthermore,} we find that underlying stretch distributions not including 
    a redshift drift are all
    excluded to describe the data as well as the Base model. In fact,
    the best non drifting model (Asymmetric) has \yc{a very marginal chance 
    ($p = 5\times10^{-6}$)} to
    describe the data as well as the Base model.}
    \yc{This result is just a quantitative assessment of qualitative facts clearly visible in 
    Fig.~\ref{fig:modelall}: the mean SN stretch per bin of redshift strongly 
    suggests a redshift evolution rather than having a constant value, and this evolution is compatible with the redshift drift expected from Eq.~\ref{eq:delta}.}

\textit{\yc{Are the next 2 paragraphs (now commented out) still pertinent? The model does not appear in table nor 
fig. Mickael, please rephrase and remove what is not needed anymore.}}

@@

\textit{\yc{This paragraph is not clear: what is the impact, and its importance, on which case? Isn't it in contradiction with the conclusions of BBC section, where it is said that proper handling of stretch correction is of paramount importance for accurate cosmology???}}
\mr{To the best of our knowledge, the SN~Ia stretch redshift drift has never
    been explicitly accounted for in cosmological analyses. Not doing so is a
    second order issue for SN cosmology as it only affects the way one does
    Malmquist bias correction. Indeed, as long as the Phillip's relation
    standardization parameter $\alpha$ is not redshift dependent (a study behind
    the scope of this paper, but see e.g. \citealt{scolnic2018a}), if a sample
    is free from selection effects, the stretch-corrected SNe~Ia magnitudes used
    for cosmology are blind to the underlying stretch distribution.  However,
    since the first missed SNe~Ia are the faintest, which are also those with
    the lowest stretch, once we include SNe~Ia affected by selection effects we
    need to account for the fact that only the brightest SNe~Ia exist in the
    low-stretch end of the sample after standardization. The modeling of the SN
    stretch then enters the analysis to understand which SNe~Ia are missing
given the selection functions to mimic a selection-bias free sample.}

@@

\mr{The state-of-the-art tool for doing so \textit{\yc{what?} \nn{mimic a selection-bias free sample}} is the BBC described in
    \cite{scolnic2016} and \cite{kessler2017}, which is used in recent SN
    cosmological analyses \citep{jones2018b, scolnic2018a, brout2019,
    descosmopaper2019} including the direct measurement of H$_0$
\citep{riess2016,riess2019}.}

\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation \yc{in our analysis}. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the smallest of all
    non-drift models (but still $\Delta \mathcal{L}=-11.5$ worse than the
    reference Base model), it is penalized by requiring 15 free parameters (3 for each of the 5 samples of the analysis).
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because \yc{traditional} surveys span limited redshift ranges, the
    per sample approach can somewhat account for \yc{implicit} redshift drifts. We
    stress here that, as measurements of modern surveys cover \yc{increasingly} larger redshift
    ranges to reduce calibration systematic uncertainties, this approach is becoming
    less \yc{valid}, notably for PS1, DES and, soon, LSST \textit{\yc{redshift ranges? Why not SNLS too?} \nn{Mickaël probably meant to cite only 3 of them?}}. Already with the current
    surveys, the BBC sample-based technique is \nn{ruled out} to be an as good
    representation of the data as our best model ($p=2\times 10^{-7}$). A more
    detailed discussion of the consequence of this results for cosmology follows
in section~\ref{sec:bbc}.}

\mr{We report in Table~\ref{tab:bbc} the samples' $\sigma_{\pm}$ and
    $\mu_0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{our} construction
    of \nn{a} selection-bias free sample, see
    section~\ref{sec:sample}. \textit{\yc{WHY??? They developed BBC for \textbf{incomplete} samples!} \nn{Yes, but their final underlying distribution
    is supposed to have been corrected for selection effects, because they were
    taken into account and added to their incomplete sample. If we ultimately find
    the same underlying distribution taking data without selection effects,
    it might mean we're doing something that concurs with what they took into
    account.}} If we were to use \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go even deeper from $-32$ to $-47$. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. \textit{\yc{???} \nn{Is the question "what's the use of this information"? If so, good question.}} \yc{Last},
    fixing the Base model parameters to those derived using SNf data only (top
    row Table~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$, \nn{and then} removing the SNf sample reduces this difference to
    $\Delta\mathrm{AIC}=-5$ \textit{\yc{not clear: the SNf sample is removed or not???} \nn{Yes. Tried to clarify with "and then".}}. Such a \yc{dramatic apparent improvement} is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift \nn{lever} arm to probe the impact of redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability $y^i$ for a SN to be young is
    known ($y^i = p_y{}^i$) while other samples rely on a probabilistic proxy for
    $y^i$, namely the fraction $\delta(z^i)$ 
    of young SNe~Ia at SN redshift. While in any case the Base model is favored, the relative weight
    of the SNf sample highlights the importance of adding non-targeted nearby
surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
graham2019} to unambiguously confirm our \nn{model}.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        distributions for the PS1 sample. The Green line (band) shows our best
        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
        approach. The dashed line shows that estimated by
    \cite{scolnic2018a}. The blue to red curves display our Base drifting
stretch model estimated at each PS SN redshift; bluer for lower redshifts.}
    \label{fig:bbc_pdf_ps1}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
    \caption{Difference of mean stretch standardization (in mag) between the BBC
        algorithm used in \cite{scolnic2018a} and our Base drifting model.
        \nn{The small colored points show the SDSS, PS1 and SNLS evolution,
            while the orange markers represent their mean evolution in 30
            regular redshift bins.} For comparison, we show in gray bands the
            expected magnitude difference between $\Lambda$CDM and $w$CDM
            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in \yc{blue},
            more exotic $w_0,w_a$CDM examples. This figure illustrates that the
            amplitude of the potential magnitude bias caused by inaccurate
        modeling of the underlying stretch distribution is comparable to the
    observational effect of exotic form of dark energy. \textit{\yc{Is this fig. still relevant?} \nn{Only as a possible representation of the potential effect of this
    difference in modeling. Why would it not anymore?}}}
    \label{fig:magdrift}
\end{figure}

@@

We have presented a study of the drift of the underlying \textit{\yc{I still prefer ``intrinsic''}} SNe~Ia stretch
distribution as a function of redshift. We used SNe~Ia from magnitude-limited
surveys from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
well as HST, to which we added SNfactory data from \cite{rigault2018} for the
low-redshift bin. We only considered the supernovae that have been discovered
in redshift ranges of each surveys where selection effect are negligible. This
way the observed SNe~Ia stretch are random sampling of the true underlying
distribution. This corresponds to 569 SNe~Ia (422 when considering more
conservative cuts).
